validateHOT - an R package for holdout task validation and market simulation

### 2024-01-04 paper/paper.Rmd
updated paper - 240104

---
title: 'validateHOT - an R package for holdout task validation and market simulation'
tags:
  - R
  - MaxDiff
  - Conjoint Analysis
  - Market Simulations
  - Predictive Validity
authors:
  - name: Joshua Benjamin Schramm
    orcid: 0000-0001-5602-4632
    corresponding: True
    affiliation: 1, 2
  - name: Marcel Lichters
    orcid: 0000-0002-3710-2292
    corresponding: FALSE
    affiliation: 2    
affiliations:
 - name: Chemnitz University of Technology, Germany
   index: 1
 - name: Otto von Guericke University of Magdeburg, Germany
   index: 2
citation_author: Schramm & Lichters
date: 02 January
year: 2024
bibliography: paper.bib
link-citations: true
output: rticles::joss_article
journal: JOSS
header-includes: 
 - \usepackage{float}
---

```{=tex}
\newcommand{\colcod}[1]{\texttt{\color{purple}#1}}
```
```{r include=FALSE}
library(validateHOT)
library(tidyverse)
```

@@

validateHOT is a package that provides functions to both validate a validation/holdout task and run market simulations for results obtained in a (adaptive) choice-based conjoint analysis (hereafter ACBC and CBC, respectively) and maximum difference scaling (hereafter MaxDiff) using, for example, `ChoiceModelR` [@ChoiceModelR] or Sawtooth's Lighthouse Studio.

@@

Preference measurement techniques', such as (A)CBC or MaxDiff, aim is to predict behavior [@green1990]. Hence, it is essential for both academics and practitioners to ensure that the collected data is valid and predicts outside tasks (i.e., the model has external validity) well.[^1] The easiest way for testing validity is by including so-called validation or holdout tasks [@Orme2015], which are tasks that are fixed (i.e., same across participants) and are usually not used for estimating the part-worth utilities in hierarchical Bayes estimation. Practitioners often do not include them [@yang2018], which is unsatisfactory given the fact that the model is used to estimate market shares which poses the basis for relevant marketing decisions.

@@

Other packages provide functions to calculate validation metrics, however, these are not specified for individual raw logit coefficients which are usually the output when running random parameter logit / hierarchical Bayes models. Metrics [@Metrics], for example, provide functions to run validation metrics such as *mean absolute error*, *root mean squared error*, or the five metrics of the confusion matrix. However, to get the output of, for example, Sawtooth Software or ChoiceModelR [@ChoiceModelR] into the right format, the user needs some data wrangling. The package conjoint [@conjoint] provides functions that are most similar to validateHOT's ones. However, no functions for validation are included and moreover, conjoint [@conjoint] mainly focuses on classical conjoint analysis. support.BWS [@support.BWS] only covers best-worst scaling case 1 (also known as Maximum Difference Scaling) and only provides market simulations based on conditional logit rule. logitr [@logitr] provides market simulations tools, however, no validation metrics such as mean hit probability [@voleti2017] or hit rate [@netzer2011]. Lastly, conjoint [@conjoint] also provides functions for both market simulations, A comparison of validateHOT's functions with current R packages is provided in \autoref{comparison}. To the best of our knowledge, a package that converts raw utility scores into validation metrics or running a variety of marketing simulations (especially TURF) is missing.

validateHOT is introduced with data estimated with Lighthouse Studio using effects-coded data for the design matrix, however, can easily be used with data estimated with ChoiceModelR [@ChoiceModelR], bayesm [@bayesm], or STAN [@rstan], if used with similar settings (ChoiceModelR, for example, automatically implements effects-coding).

```{=tex}
\begin{figure}[h]
  \includegraphics{figures/FunctionComparison.png}
  \caption{Comparison of validateHOT's function to existing R packages}
  \label{comparison}
\end{figure}
```
# Key functions

validateHOT's functions can be categorized into four main areas, see \autoref{tab:table1}. To bring the data into the right format, users can run the \texttt{\color{purple}createHOT()} function, which creates the total utility of each alternative by applying the additive utility model [@rao2014, p. 82]. \colcod{turf()} as well as the four rescaling functions, however, are not dependent on \texttt{\color{purple}createHOT()}, and can be run using the raw logit scores.

@@

: Overview of main four areas of validateHOT and their corresponding functions \label{tab:table1}

@@

After running the hierarchical Bayes estimation [@allenby1995; @lenk1996], the **raw** utility scores have to be exported and read into an *R* data frame. This data frame has to include the actual choice in the validation/holdout task.

@@

@@

Finally, we show two functions for market simulations, namely \texttt{\color{purple}marksim()} and \texttt{\color{purple}turf()}. In the following example, the market share is calculated according to the multinomial logit model [@McFadden1974].

@@

Finally, \texttt{\color{purple}turf()}, a "product line extension model" [@miaoulis1990, p. 29], is a tool to find the perfect assortment that creates the highest reach and is especially powerful for MaxDiff studies [@chrzan2019, p. 108]. To optimize the search for the optimal bundle, we also include the arguments \texttt{\color{purple}fixed}, to define alternatives that have to be part of the assortment, and \texttt{\color{purple}prohib}, to prohibit certain item combinations of being part of the assortment (see the vignette for more details and how to apply \texttt{\color{purple}turf()} with data obtained using a likert scale).

@@

Again, we start by creating the validation scenario.

@@

This time we calculate the mean hit probability.

@@

Finally, we can also display the attribute importance scores. Therefore, we need to define the attribute levels as well as the coding of the attributes.

>>>>

---
title: 'validateHOT - an R package for holdout task validation and market simulation'
tags:
  - R
  - MaxDiff
  - Conjoint Analysis
  - Market Simulations
  - Predictive Validity
authors:
  - name: Joshua Benjamin Schramm
    orcid: 0000-0001-5602-4632
    corresponding: True
    affiliation: 1, 2
  - name: Marcel Lichters
    orcid: 0000-0002-3710-2292
    corresponding: FALSE
    affiliation: 2    
affiliations:
 - name: Chemnitz University of Technology, Germany
   index: 1
 - name: Otto von Guericke University of Magdeburg, Germany
   index: 2
citation_author: Schramm & Lichters
date: 02 January
year: 2024
bibliography: paper.bib
link-citations: true
output: rticles::joss_article
journal: JOSS
header-includes: 
 - \usepackage{float}
 - \usepackage{fancyhdr}
---

```{=tex}
\setlength{\headheight}{63.55022pt}
\addtolength{\topmargin}{-0.95425pt}
\newcommand{\colcod}[1]{\texttt{\color{purple}#1}}
```

```{r include=FALSE}
library(validateHOT)
library(tidyverse)
```

@@

validateHOT is an R package that provides functions to both validate a validation/holdout task and run market simulations for results obtained in a (adaptive) choice-based conjoint analysis (hereafter ACBC and CBC, respectively) and maximum difference scaling (hereafter MaxDiff) using, for example, ChoiceModelR [@ChoiceModelR] or Sawtooth's Lighthouse Studio.

@@

Preference measurement techniques' (e.g., (A)CBC or MaxDiff) aim is to predict behavior [@green1990]. Hence, it is essential for both academics and practitioners to ensure that the collected data is valid and predicts outside tasks (i.e., the model has external validity) well.[^1] The easiest way for testing validity is by including so-called validation or holdout tasks [e.g., @rao2014; @Orme2015], which are tasks that are fixed (i.e., same across participants) and are usually not used for estimating the part-worth utilities in hierarchical Bayes estimation. Practitioners often do not include them [@yang2018], which is unsatisfactory given the fact that the model is used to estimate market shares which poses the basis for relevant marketing decisions.

@@

Other packages provide functions to calculate validation metrics, however, these are not specified for individual raw logit coefficients which are usually the output when running random parameter logit / hierarchical Bayes models. Metrics [@Metrics], for example, provide functions to run validation metrics such as *mean absolute error*, *root mean squared error*, or the five metrics of the confusion matrix. However, to get the output of, for example, Sawtooth Software or ChoiceModelR [@ChoiceModelR] into the right format, the user needs some data wrangling. The package conjoint [@conjoint] provides functions that are most similar to validateHOT's ones. However, no functions for validation are included and moreover, conjoint [@conjoint] focuses on classical conjoint analysis, and thus is limited when applying more common CBC methods, for example, (A)CBC. support.BWS [@support.BWS] only covers best-worst scaling case 1 (also known as MaxDiff) and only provides market simulations based on conditional logit rule. logitr [@logitr] provides market simulations tools, however, no validation metrics such as mean hit probability [@voleti2017] or hit rate [@netzer2011]. A comparison of validateHOT's functions with current R packages is provided in \autoref{comparison}. To the best of our knowledge, a package that converts raw utility scores into validation metrics or running a variety of marketing simulations (especially TURF) is missing.

```{=tex}
\begin{figure}[h]
  \includegraphics{figures/FunctionComparison.png}
  \caption{Comparison of validateHOT's function to existing R packages}
  \label{comparison}
\end{figure}
```

validateHOT is introduced with data estimated with Lighthouse Studio using effects-coding for creating the design matrix. It, however, can easily be used with data estimated with ChoiceModelR [@ChoiceModelR], bayesm [@bayesm], or STAN [@rstan], if used with similar settings (ChoiceModelR, for example, automatically implements effects-coding).

# Key functions

validateHOT's functions can be categorized into four main components, see \autoref{tab:table1}. To bring the data into the right format, users can run the \texttt{\color{purple}createHOT()} function, which creates the total utility of each alternative by applying the additive utility model [@rao2014, p. 82]. \colcod{turf()} as well as the four rescaling functions, however, are not dependent on \texttt{\color{purple}createHOT()}, and can be run using the raw logit scores.

@@

: Overview of validateHOT's main four components and their corresponding functions \label{tab:table1}

@@

### Creating Holdout Task / Market Scenario

After running the hierarchical Bayes estimation [@allenby1995; @lenk1996], the **raw** utility scores have to be exported and read into an *R* data frame. This data frame has to include the actual choice in the validation/holdout task (if only a market scenario is created, the \texttt{\color{purple}choice} argument in \texttt{\color{purple}createHOT()} can be left empty)..

@@

### Validating Holdout Task

@@

### Market Simulations

Lastly, we introduce two functions for market simulations, namely \texttt{\color{purple}marksim()} and \texttt{\color{purple}turf()}. In the following example, the market share is calculated according to the multinomial logit model [@McFadden1974].

@@

Next, \texttt{\color{purple}turf()}, a "product line extension model" [@miaoulis1990, p. 29], is a tool to find the perfect assortment that creates the highest reach and is especially powerful for MaxDiff studies [@chrzan2019, p. 108]. To optimize the search for the optimal bundle, we also include the arguments \texttt{\color{purple}fixed}, to define alternatives that have to be part of the assortment, and \texttt{\color{purple}prohib}, to prohibit certain item combinations of being part of the assortment (see the vignette for more details and how to apply \texttt{\color{purple}turf()} with data obtained using a likert scale).

@@

### Creating Holdout Task / Market Scenario

The setup is almost the same, only the arguments \texttt{\color{purple}prod.levels}, \texttt{\color{purple}coding}, and \texttt{\color{purple}method} are different or new, respectively.
\newpage

@@

### Validating Holdout Task

This time we calculate the mean hit probability (i.e., MHP).

@@

### Rescaling Scores

Finally, we can also display the attributes importance scores. Therefore, we need to define the attribute levels as well as the coding of the attributes.
